### 2013-05-19 docs/replication_analysis/analyse_replication.Rnw
analysing replication results

\documentclass[paper=a4, fontsize=11pt]{scrartcl}
\title{Analysis of initial replication results}

\begin{document}
\maketitle

\section{First look}

<<read_summary, eval=TRUE, echo=FALSE, warning=FALSE, cache=TRUE>>=
source("~/repo/eQTL-2D/replication/results/summary_script.R")
@

\end{document}