### 2013-09-09 epistasis_pages/index.Rmd
Supporting web page to be hosted on github pages

---
title: Epistasis Detection and Replication
subtitle: 
author: Gibran Hemani, Konstantin Shakhbazov
github: {user: , repo: , branch: }
framework: minimal
mode: selfcontained
widgets: [polycharts]
highlighter: highlight.js
hitheme: solarized_light
assets:
  css: 
    - "http://fonts.googleapis.com/css?family=Open+Sans"
    - "http://fonts.googleapis.com/css?family=Open+Sans+Condensed:700"
---

# Epistasis Detection adn Replication

This is web page of a paper [Detection and replication of epistasis influencing transcription in humans](http://www.nature.com)

# Abstract

Epistasis is the phenomenon whereby one polymorphism’s effect on a trait depends on other polymorphisms present in the genome. The ex- tent to which epistasis influences complex traits1 and contributes to their variation2,3 is a fundamental question in evolution and human genetics. Though epistasis has been demonstrated in artificial gene manipulation studies in model organisms,4,5 and some examples have been reported in other species,6 few convincing examples exist for epistasis amongst natural polymorphisms in human traits.7,8 Its absence from empirical findings may simply be due to its low incidence in the genetic control of complex traits,2,3 but an alternative view is that it has previously been too technically challenging to detect due to statistical power and computational issues.9 Here we show that, using advanced computation techniques10 and a gene expression study design, many instances of epis- tasis are found between common single nucleotide polymorphisms (SNPs). In a cohort of 846 individuals with data on 7339 gene expression levels in peripheral blood, we found 501 significant pairwise epistatic interac- tions between common SNPs acting on the expression levels of 238 genes (p < 2.91 × 10−16). We tested the discovery interactions for replication in two independent data sets.11,12 Three hundred and forty-five interac- tions had replication interaction p-values that were more extreme than the 2.5% confidence interval of the distribution under the null hypothesis of no epistasis, with 30 significant at a conservative p < 0.05 Bonferroni level. There was evidence of functional enrichment for the interacting SNPs, for instance 44 of the genetic interactions are located within 2Mb of re- gions of known intra-cellular chromosome interactions13 (p = 1.8×10−10). Epistatic networks of three SNPs or more influence the expression levels of 129 genes, whereby one cis-acting SNP is modulated by several trans- acting SNPs. For example MBNL1 is influenced by an additive effect at rs13069559 which itself is masked by trans-SNPs on 14 different chro- mosomes, with nearly identical genotype-phenotype (GP) maps for each cis-trans interaction. This study presents the first evidence for multiple instances of epistatic genetic effects emerging from natural genetic varia- tion in humans.